%
%  Abstract
%

\begin{abstract}
\addcontentsline{toc}{chapter}{Abstract}

Cloud computing has revolutionized the way developers implement and deploy applications.
By running applications on large-scale compute infrastructures and programming
platforms that are remotely accessible as utility services, cloud computing provides scalability, 
high-availability, and increased user productivity.

Despite the advantages inherent to the cloud computing model, it has also given rise to several
software management and maintenance issues. Specifically, cloud platforms do not
enforce developer best practices, and other administrative requirements when 
deploying applications. Cloud platforms also do not facilitate
establishing service level objectives (SLOs) on application performance, which are
necessary to ensure reliable and consistent operation of applications.
Moreover, cloud platforms do not provide adequate support to monitor the performance
of deployed applications, and conduct root cause analysis when an application exhibits a
performance anomaly.

We employ \textit{governance} as a methodology to address the above mentioned issues prevalent in
cloud platforms. We devise novel governance solutions that
achieve administrative conformance, developer best practices, and performance SLOs in the cloud via
policy enforcement, SLO prediction, performance anomaly detection and root cause analysis.
The proposed solutions are fully automated, and built into the cloud platforms as cloud-native features thereby
precluding the application developers from having to implement similar features by themselves. 
We evaluate our methodology using real world cloud platforms, and show that
our solutions are highly effective and efficient.

%\abstractsignature
\end{abstract}
