\begin{vitae}
\addcontentsline{toc}{chapter}{Curriculum Vitae}

\begin{vitaesection}{Education}
\vspace{-0.1cm}
\item [2016]	Ph.D. in Computer Science (Expected), \\ University of California, Santa Barbara, United States.
\item [2009]	B.Sc. Engineering (Hons) Degree, \\ University of Moratuwa, Sri Lanka.
\end{vitaesection}

\textbf{Publications}

\textit{Service-Level Agreement Durability for Web Service Response Time} \\
H. Jayathilaka, C. Krintz, R. Wolski \\
International Conference on Cloud Computing Technology and Science (CloudCom), 2015. \\

\textit{Response time service level agreements for cloud-hosted web applications} \\
H. Jayathilaka, C. Krintz, and R. Wolski \\
ACM Symposium on Cloud Computing (SoCC), 2015. \\

\textit{EAGER: Deployment-Time API Governance for Modern PaaS Clouds} \\
H. Jayathilaka, C. Krintz, and R. Wolski \\
IC2E Workshop on the Future of PaaS, 2015. \\

\textit{Using Syntactic and Semantic Similarity of Web APIs to Estimate Porting Effort} \\
H. Jayathilaka, A. Pucher, C. Krintz, and R. Wolski \\
International Journal of Services Computing (IJSC), 2014. \\

\textit{Towards Automatically Estimating Porting Effort between Web Service APIs} \\
H. Jayathilaka, C. Krintz, and R. Wolski \\
International Conference on Services Computing (SCC), 2014. \\

\textit{Cloud Platform Support for API Governance} \\
C. Krintz, H. Jayathilaka, S. Dimopoulos, A. Pucher, R. Wolski, and T. Bultan \\
IC2E Workshop on the Future of PaaS, 2014. \\

\textit{Service-driven Computing with APIs: Concepts, Frameworks and Emerging Trends} \\
H. Jayathilaka, C. Krintz, and R. Wolski \\
IGI Global Handbook of Research on Architectural Trends in Service-driven Computing, 2014. \\

\textit{Improved Server Architecture for Highly Efficient Message Mediation} \\
H. Jayathilaka, P. Fernando, P. Fremantle, K. Indrasiri, D. Abeyruwan, S. Kamburugamuwa, S. Jayasumana, S. Weerawarana and S. Perera \\
International Conference on Information Integration and Web-based Applications and Services (IIWAS), 2013. \\

\textit{Extending Modern PaaS Clouds with BSP to Execute Legacy MPI Applications} \\
H. Jayathilaka and M. Agun \\
ACM Symposium on Cloud Computing (SoCC), 2013. \\

\end{vitae}