Cloud computing has revolutionized the way software is
developed, deployed, and consumed.  As a consequence, there has been a
proliferation of cloud-hosted applications, which developers make accessible to users via
web application programming interfaces (web APIs).
Because this model significantly simplifies and expedites
deployment of web APIs, it also poses new software maintenance and evolution
challenges.  In particular, it becomes difficult to track, control, and compel
reuse of web APIs, inadequately tested applications can be
deployed into production, and API changes can be introduced that
break API-user code or that breach security or organizational procedures.

To address these challenges, we investigate a new approach to API governance
in the cloud -- combined policy, implementation, and deployment control of APIs for
software and data deployed as cloud-hosted applications.  Our approach, called EAGER,
provides a software architecture that can be easily integrated into cloud
platforms as a cloud-native feature, and supports system-wide, 
deployment-time enforcement of API governance policies.
Specifically, EAGER can check for and
prevent backward incompatible API changes from
being deployed into production,
enforces service reuse, and facilitates enforcement of other best practices
in software maintenance via policies.  We also describe a prototype
EAGER implementation that
integrates with an open source platform-as-a-service cloud, and
evaluate its feasibility, efficiency,
scalability, and effectiveness for enforcing cloud-based API governance.

