As the paradigm of cloud computing grows in popularity, the need for monitoring cloud-hosted 
applications is becoming critical. Application developers and cloud administrators
wish to detect performance anomalies in cloud applications, and
perform root cause analysis to diagnose problems. Uncovering performance bottlenecks in a timely manner,
and resolving them urgently is also a key requirement for implementing governance in cloud
environments. However, the high level of abstraction provided by cloud
platforms, coupled with their scale and complexity, makes performance diagnosis
a daunting problem. This situation is particularly apparent in
PaaS clouds, where the application runtime details are hidden beneath a layer of curated services.
The existing cloud monitoring solutions do not have the necessary penetrative power
to monitor all the different layers of such cloud platforms, and consequently, their diagnosis
capabilities are severely limited.

We propose Roots, a near real time monitoring framework for applications deployed in a PaaS cloud. 
Roots is designed to function as a curated service
built into the cloud platform, as opposed to an external monitoring system. 
It relieves the application developers from having to configure
their own monitoring solutions, or having to instrument the application code in anyway.
Roots captures runtime data from all the different layers involved
in processing application requests. It can correlate events across different layers, and
identify bottlenecks deep within the kernel services of the PaaS.

Roots monitors applications for SLO compliance, and detects anomalies via SLO violations.
When Roots detects an anomaly, 
it analyzes workload data and other application runtime data
to perform root cause analysis. Roots is able to determine whether a particular
anomaly was caused by a change in the application workload, or due to a bottleneck
in the cloud platform. To this end we also devise a bottleneck identification algorithm, that
uses a combination of linear regression, quantile analysis and change point detection. 
We also present an analysis method by which Roots can identify different paths of execution
in an application. Our method does not require static analysis, and we use it to detect
changes in an application's workload characteristics.

We evaluate Roots using a prototype built for the AppScale open source PaaS. 
Our results indicate that Roots is effective at detecting performance anomalies
in near real time. We also show that our bottleneck identification algorithm
produces accurate results nearly 100\% of the time, pinpointing the exact PaaS kernel
service or the application component responsible for each anomaly. Our empirical trials further 
reveal that Roots does not add a significant overhead to the applications deployed
on the cloud platform. Finally, we show that Roots is very lightweight, 
and scales well to handle large populations of applications.

In our future work we plan to expand the data gathering capabilities of Roots into
the low level virtual machines, and containers that host various services of the cloud
platform. We intend to tap into the hypervisors
and container managers to harvest runtime data regarding the resource usage (CPU, memory, disk etc.) of
PaaS services and other application components. With that we expect to extend
the root cause analysis support of Roots so that it can not only pinpoint the
bottlenecked application components, but also the low level hosts and system
resources that constitute each bottleneck.