We present Roots, a full-stack monitoring and analysis system
for performance anomaly detection and bottleneck identification
in cloud platform-as-a-service (PaaS) systems. Roots facilitates
application performance monitoring as a core capability of PaaS clouds, and
relieves the developers from having to instrument application code. Roots tracks 
HTTP/S requests to hosted cloud applications and their use of PaaS services.
To do so it employs lightweight monitoring of PaaS service interfaces.
Roots processes this data in the background using multiple
statistical techniques that in combination detect performance
anomalies (i.e. violations of service-level objectives). For each anomaly,
Roots determines whether the event was caused by a change in
the request workload or by a performance bottleneck in a PaaS service. By
correlating data collected across different layers of the PaaS, Roots is able
to trace high-level performance anomalies to bottlenecks in specific components
in the cloud platform.
We implement Roots using the AppScale PaaS and evaluate its overhead
and accuracy.

%Roots performs lightweight, full stack monitoring by integrating
%the APM as a managed PaaS service.  By doing so, Roots is able to 
%efficiently collect and correlate performance data across the PaaS stack,
%precluding the need for application instrumentation.  Roots employs
%a combination of statistical techniques that analyze this data to detect performance
%anomalies (i.e. violations of service-level objectives). For each anomaly, 
%Roots determines whether the event was caused by a change in 
%the workload, and if not identifies the bottleneck component in the cloud platform
%that is most likely to have degraded application performance. 
%We implement Roots in the AppScale open source cloud platform, and evaluate its 
%accuracy and overhead using PaaS-hosted web applications.

