Cloud computing delivers IT infrastructure resources, programming platforms, and software
applications as shared utility services. Enterprises and developers increasingly deploy applications 
on cloud platforms due to the scalability, high availability and many other
productivity enhancing features of the cloud. Cloud-hosted applications 
depend on the core services provided
by the cloud platform for resources. In some cases they use the services provided by the cloud to implement most of
the application functionality as well (e.g. PaaS-hosted applications). 
Cloud-hosted applications are typically
accessed over the Internet, via the web APIs exposed by the applications.

As the applications hosted in cloud platforms continue to increase in number, the need for enforcing
governance on them becomes accentuated. We define governance as the mechanism by which the 
acceptable operational parameters are specified and maintained for a cloud-hosted application.
Governance enables specifying the acceptable
development standards and runtime parameters (performance, availability, security requirements etc.) 
for cloud-hosted applications as policies. Such
policies can then be enforced automatically at various stages of the application life-cycle. 
Governance further entails
monitoring cloud-hosted applications to ensure that they operate at a certain level of quality,
and taking corrective action when problems are detected. Through the steps of specification,
enforcement, monitoring and correction, governance can help resolve a number of prevalent issues in
today's cloud platforms. These issues include lack of good software engineering practices (code reuse,
dependency management, versioning etc), lack of performance SLAs for cloud-hosted applications,
and lack of performance debugging support. 

This proposal is aimed at exploring the feasibility of efficiently enforcing governance on cloud-hosted
applications, and evaluating the effectiveness of governance as a means of achieving administrative
conformance, developer best practices and performance objectives in the cloud. Considering the scale of
today's cloud platforms in terms of the number of users and the applications, 
we wish to automate much of the governance tasks through
automated analysis and diagnostics. To achieve efficiency, we put more emphasis on deployment-time
policy enforcement and non-invasive passive monitoring of cloud platforms, thereby keeping the 
governance overhead
to a minimum. We avoid run-time enforcement and invasive instrumentation of cloud applications 
as much as possible. We also focus on building governance systems that are deeply integrated with
the cloud platforms themselves. This enables using the existing scalability and high availability features of the cloud
to provide an efficient governance solution that can control all application events in a fine-grained
manner. Furthermore, such integrated solutions relieve the users from having to maintain and pay
for additional, external governance and monitoring solutions.

With the goal of efficient, integrated and automated governance for cloud-hosted applications in mind,
we propose the following three research goals.
\begin{enumerate}
\item Design and implement a scalable, low-overhead governance framework for cloud platforms,
complete with a policy specification language and a policy engine. The governance framework should be
integrated into the cloud platforms, and must
strive to keep the runtime overhead of the user applications to a minimum while enforcing
common developer best practices and other organizational conventions.
\item Design and implement a methodology for formulating performance SLAs for cloud-hosted 
 web applications, without
 subjecting them to extensive performance testing or instrumentation. The formulated
 performance SLAs must be correct, tight and durable in the face of changing
 conditions of the cloud.
 \item Design and implement a scalable cloud application monitoring framework for detecting
application performance anomalies, SLA violations and diagnosing potential root causes. 
The framework should support collecting
 a wide range of performance and usage data from the user applications and the underlying cloud platform
 without instrumenting the user code, and without introducing significant runtime overheads.
 It should further facilitate employing a number of new and existing statistical methods
 and algorithms to analyze the gathered data in near realtime.
\end{enumerate}

We design and implement EAGER~\cite{eager-fop15} as a means of achieving the first goal. EAGER is a lightweight
governance framework built into PaaS clouds. It supports defining policies using a simple syntax
based on the popular Python programming language. EAGER promotes deployment-time
policy enforcement, where policies are enforced on user applications (and APIs) every time
an application is uploaded to the cloud. By carrying out policy validations at
application deployment-time, and refusing to deploy applications that violate policies,
 we provide fail-fast semantics which ensure that deployed applications are fully policy compliant. 
EAGER architecture also provides the necessary provisions for facilitating run-time policy
enforcement (through an API gateway proxy) if necessary. Our experimental results show
that EAGER validation and policy enforcement overhead is very small, and it scales well to
handle thousands of user applications and policies. Overall, we show that integrated governance
for cloud-hosted applications is not only feasible, but also can be implemented with very
little overhead and effort.