Increasingly, web, mobile, and cloud developers integrate
web services 
exposed via well-defined APIs into their Internet accessible applications.  
Doing so simplifies and expedites software development, testing, 
deployment, and management of these applications. As a result, there has been 
a proliferation of web APIs, and this model
of software development has broadened participation in computing.
In particular, web APIs have become ubiquitous and influence every 
aspect of society including commerce, entertainment, communication, 
health care and security.  

As cloud computing continues to garner widespread adoption, the number 
of web APIs served from cloud platforms has begun to explode. But
today's cloud platforms offer little or no support for enforcing API governance
(i.e. policy compliance at API development time and runtime). 
Also there are no straightforward mechanisms
for analyzing the performance and QoS properties of APIs hosted in clouds
without having to
perform a lot of offline and/or online testing. As a result developers today have
no easy way of understanding the SLAs that can be supported from their services or
a way to enforce them under pre-calculated confidence levels.

By providing simplified, unified,
and automated governance and SLA enforcement for cloud-based 
web APIs, my work has the potential 
to reduce the burden on IT operations associated with 
API development and management, and to improve the robustness of real systems for
organizations across all sectors of our society.